MLH1 Results Outline

### 2020-02-12 doc/Results.rmd
more updates and knit

```{r mixed.model.data.setup, echo=FALSE, warning=FALSE, eval=FALSE}
#DF.HetC.MixedModel.HQ

@@

*Do I believe the physical/SC results?*

@@

3) Interference / IFD will not be sexually dimorphic.

B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.

**I**nter**f**ocal **D**istance, a indication of CO interference.

The raw measures are pretty nosy, this might not be the best way to display these results.

 Each point is a IFD observation. Maybe add mouse level lines for the means.

@@

### IFD_HetC_M1

```{r HetC_M1_IFD_many.metrics, echo=FALSE}
#get a list of the old mice to double check
#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
#MIXED MODEL

@@

@@

#these 2 below are LR tests
#drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
#anova(lmer.IFD.PER_MM.M1)

#exactRLRT(lmer.IFD.PER_MM.M1)
#as.data.frame(round(summary(lmer.IFD.PER_MM.M1)$coefficients, 6))

#M1_IFD.RAW_2CO
lmer.IFD.RAW_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

#these 2 below are LR tests
#drop1(lmer.IFD.RAW_MM.M1, test="Chisq" ) #other options for test don't work
#anova(lmer.IFD.RAW_MM.M1) # also F values

#exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
#as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
```

Sex is a slightly significant factor for raw and normalized IFD.

@@

show a raw plot, and norm plot and a scatter plot

@@

### IFD_HetC_M2

@@

### IFD_HetC_M3

@@

@@

```{r CO.positions.figs, echo=FALSE, include=FALSE}
#main figure of density plots,

@@

```{r show.long.biv, echo=FALSE}
#real.long.bivData

@@

@@

Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
#summary(Q1.glm_nrmF1pos_M2)

@@

```{r Q1.Nrm.F1pos_M2_long.biv, echo=FALSE}

#table for the 1CO long biv across strain and sexes?

@@

##### Mixed models norm 1CO positions

@@

#### Nrm_1COpos_HetC_M1

```{r Q1.MM_CO_pos, echo=FALSE, eval=FALSE}
#use the 1CO
#Use MELT DF
#MELT_Curated_BivData

#isolate the 1COPos.CO.avs
#Pos.CO.avs_1CO <- Pos.CO.avs[Pos.CO.avs$Foci_Type == "Foci1",]
#Pos.CO.avs_1CO <- Pos.CO.avs_1CO[!(is.na(Pos.CO.avs_1CO$mouse) | Pos.CO.avs_1CO$mouse==""), ]
#BivData.CO.PER.POS.lme <- lme(mean_PER_pos ~ subsp * sex, data=Pos.CO.avs_1CO, random=list(strain=pdDiag(~sex) ) )
#as.data.frame( round(  summary(BivData.CO.PER.POS.lme)$tTable, 6 ) )

mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]

#<- lme()
lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )

#can't make the 
CO_position.MM <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)

#library(lmerTest)
#anova(CO_position.MM)
#anova(CO_position.MM, verbose = TRUE)

#summary(CO_position.MM) #no p values,,
#these just give t and F stats, no pvalue

#CO_pos_fix.est <- drop1(CO_position.MM, test="Chisq")
#CO_pos_fix.est
#this test, the full model is more predictive than an empty model?

#as.data.frame( round(  summary(CO_position.MM)$tTable, 6 ) )  #lme() needed for this
#thing2 <- drop1(SC_length.MM, test="Chisq")

CO_pos_rand <- exactRLRT(CO_position.MM)
CO_pos_rand
```

```{r Q1.nrmFoci1_M1, echo=FALSE}

#M1

@@

#### Nrm_1COpos_HetC_M2

@@

```{r var.contrast, echo=FALSE, eval=FALSE}
#contrast models -- what's the points for this block
# review this logic again
full2strain.CV <- anova(HetC.MixedModel.varco, just.strain.var_co)#0.9997

full2noRandom.CV <- anova(HetC.MixedModel.varco, full.Reduced.var_co) #0.3012

strain2noRandom.CV <- anova(just.strain.var_co, full.Reduced.var_co)#0.1213
#for var, p = 0.0219
```

>>>>

```{r mixed.model.data.setup, echo=FALSE, warning=FALSE, eval=TRUE}
#DF.HetC.MixedModel.HQ

@@

@@

@@

```{r Q1.IFD.M1, echo=FALSE}
#get a list of the old mice to double check
#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
#MIXED MODEL

@@

#M1_IFD_2CO
lmer.IFD.ABS_MM.M1 <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

lmer.IFD.ABS_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_ABS ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )

lmer.IFD.ABS_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_ABS ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )

anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.sex)
anova(lmer.IFD.ABS_MM.M1, lmer.IFD.ABS_MM.M1.reduce.subp)

drop1(lmer.IFD.ABS_MM.M1, test="Chisq" )

exactRLRT(lmer.IFD.ABS_MM.M1)

@@

lmer.IFD.PER_MM.M1.reduce.sex <- lmer( mean_IFD.2CO_PER ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )

lmer.IFD.PER_MM.M1.reduce.subp <- lmer( mean_IFD.2CO_PER ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )

anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.sex)
anova(lmer.IFD.PER_MM.M1, lmer.IFD.PER_MM.M1.reduce.subp)

#these 2 below are LR tests
drop1(lmer.IFD.PER_MM.M1, test="Chisq" )
exactRLRT(lmer.IFD.PER_MM.M1)

@@

@@

$$mouse \ average \ IFD ~=~ subsp * sex * strain + \varepsilon $$

@@

glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)

summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
#round(summry.glm.Q1_IFD.ABS.M2$coefficients[,c(1,4)],5)

#write.table(round(summry.glm.Q1_IFD.ABS.M2$coefficients,5), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M2.csv", sep=",",row.names = TRUE)

#PER
glm.Q1_IFD.PER.M2 <- glm(mean_IFD.2CO_PER ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)

summry.glm.Q1_IFD.PER.M2 <- summary(glm.Q1_IFD.PER.M2)
#round(summry.glm.Q1_IFD.PER.M2$coefficients[,c(1,4)],5)

#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)

$$mouse \ average \ IFD ~=~  sex * strain + \varepsilon $$

@@

glm.Q1_IFD.ABS.M3  <- glm(mean_IFD.2CO_ABS ~  sex * strain, data=Mouse.Table_BivData_4MM)

summry.glm.Q1_IFD.ABS.M3 <- summary(glm.Q1_IFD.ABS.M3)
round(summry.glm.Q1_IFD.ABS.M3$coefficients[,c(1,4)],5)

write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)

#PER
glm.Q1_IFD.PER.M3  <- glm(mean_IFD.2CO_PER ~ sex * strain, data=Mouse.Table_BivData_4MM)
summry.glm.Q1_IFD.PER.M3 <- summary(glm.Q1_IFD.PER.M3)
round(summry.glm.Q1_IFD.PER.M3$coefficients[,c(1,4)],5)

write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)

For the Mixed models of IFDs, sex is a significant effect for both raw and nrmIFD. for the nrm.IFD, subspecies.

the interaction effects were slightly significant for both raw and nrm.IFD.

The most significant value was the sex effect for nrm.IFDs.

The the random strain effect was not significant for either model.

For the raw measures, in M2, I think the 2 SKIVE effects mean that the female raw IFD is shorter than the male IFD.raw. The other effects are for PWD ansD SKIVE (larger raw IFD from intercept.)  I think the MSM and MOLF interaction effects were too far down the list to sop up any variance. For the M3, only SKIVE*male effect is close to significant

For the normalized values in both M2 and M3, sex is a significant effect, increasing nrm.IFD in males. SKIVE*male is the only other consistantly significant effect, which also increases the nrm.IFD measure.

Overall There's a low amount of significant effects across the 2CO IFD measures. This might be an indication that interference is conserved across these samples and/or that there is too much noise across from chromosome specific effects.

@@

density plots
```{r CO.positions.figs, echo=FALSE, include=FALSE}
#main figure of density plots,

@@

@@

```{r Q1.nrmF1CO.pos_M1, echo=FALSE}

lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )

lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )

anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)

drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )

exactRLRT(lmer.Q1.nrm1CO_M1)

#write.table(round(summry.glm.Q1_IFD.PER.M2$coefficients, 6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M2.csv", sep=",",row.names = TRUE)

$$mouse \ average \ F1 position ~=~ subsp * sex * strain + \varepsilon $$

@@

Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Mouse.Table_BivData_4MM)
#summary(Q1.glm_nrmF1pos_M2)

summry.Q1.glm_nrmF1pos_M2 <- summary(Q1.glm_nrmF1pos_M2)
round(summry.Q1.glm_nrmF1pos_M2$coefficients[,c(1,4)],5)

write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", sep=",",row.names = TRUE)

@@

$$mouse \ average \ F1 position ~=~ sex * strain + \varepsilon $$

```{r Q1.nrmCO.pos_M3, echo=FALSE, eval=FALSE}

Q1.glm_nrmF1pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
#summary(Q1.glm_nrmF1pos_M2)

summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)

write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
```

The model results aren't as clear as I'd like.
For the Mol, both the sex effect and MOLF strain effect are significant (this means both female and male rec landscapes are effected - most of the time towards the middle

GENERALLY - Sex is the biggest effect for the 1CO landscape (female middle male telomereic)
molf
Not sure I understand how MSM is supper significant for 
There must be some strange effect due to the Mol subsp - MOLF isn't even in M2 (
Male is consistently significant across the two models.

Male and MOLF are the most significant effects for M3

@@

@@

@@

@@
